---
title: "Reef clustering - Post processing"
output: pdf_document
---

This document shows how to post-process the Dirichlet mixture MCMC chains using the functions in \verb|chain_processing.R|. Our first chunck of code checks if the required libraries are available, installing them if necessary. Afterwards, function in \verb|chain_processing.R| are loaded.

```{r, warning=FALSE, message=FALSE}
packages <- c("ggplot2", "dplyr", "data.table", "reticulate", 
              "label.switching", "abind", "tidyr", 'plyr')

for(p in packages){
  if(!require(p,character.only = TRUE)) install.packages(p)
  library(p,character.only = TRUE)
}

source('chain_processing.R')
```

After ensuring that the MCMC file is in our working directory, we can use \verb|read_mcmc_file| to read it into R.

```{r, eval=FALSE}
filename= "Example/MCMC_example.p"
results = read_mcmc_file(filename)
```

Label switching is dealt with by using \verb|relabel_mcmc| which applies Stephen's relabelling algorithm, requiring the package \verb|label.switching|. This returns a list with relabelled $\boldsymbol{\underline{\rho}}$ and $\boldsymbol{z}$ parameters.

```{r, eval = FALSE}
results_relab = relabel_mcmc(results)
```

We can now obtain summaries of interest of the parameters and latent variables posterior distributions. The following chunck of code gathers the 5%, 50% and 95% quantiles of the $\boldsymbol{\underline{\rho}}$ elements.

```{r, message=FALSE, warning=FALSE}
results_relab$rho %>%
  group_by(group, prop) %>%
  dplyr::summarise( Q5 = quantile(rho, probs = 0.05),
                    median = median(rho, na.rm = T),
                    Q95 = quantile(rho, probs = 0.95))
```

An auxiliary function \verb|plot_rho_posterior| is also available to easily obtain the trace by chains and density plots of the combined draws. 

```{r, fig.width=8, fig.height=5, fig.align='center'}
rho_plots = plot_rho_posterior(results_relab)
rho_plots$trace
```

```{r, fig.width=8, fig.height=5, fig.align='center'}
rho_plots$density
```

Finally, we can summarise the cluster allocation of the observations with \verb|summary_z|. This function returns the posterior cluster allocation probability for each observation in the sample as well as the mode partition.

```{r}
z_post = summary_z(results_relab$z, k= 2)
head(z_post$probability)

z_post$mode
```



